\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

%TODO: Acknowledgments

I am profoundly grateful to my advisor, Samuel Torre Vinuesa, whose steadfast guidance and insightful criticism have been instrumental in my journey through this research. His deep knowledge and clear articulation of the subject matter have significantly enhanced my understanding and appreciation of the field. Samuel's support in defining the research topic, structuring the thesis, and navigating the challenges encountered during my research journey has been invaluable. 

Additionally, I extend my heartfelt thanks to my supervisor, Prof. Dr.-Ing. Michael Gerndt, for offering me the chance to conduct my research within the chair of Computer Architecture and Parallel Systems at the Department of Informatics, Technical University of Munich. This opportunity has been pivotal in my academic and professional growth. 

My colleagues at Amadeus Data Processing GmbH deserve special mention for their unwavering support throughout my research. Jerome Mazeries and Gais Ameer's assistance with technical issues and their constructive feedback have been crucial in the completion of my work. I am deeply appreciative of the supportive and collaborative environment they provided. 

Last but certainly not least, I owe a tremendous debt of gratitude to my family. Their moral support and belief in my abilities have been my stronghold. It is their encouragement that fortified my resolve and enabled me to bring this thesis to fruition. 

\cleardoublepage{}
